---
title: "Getting SEC DERA XBRL data"
---

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The SEC publishes two bulk XBRL data sets that are useful for financial
statement research:

- [Financial Statement Data Sets](https://www.sec.gov/data-research/sec-markets-data/financial-statement-data-sets)
- [Financial Statement and Notes Data Sets](https://www.sec.gov/data-research/financial-statement-notes-data-sets)

The `dera.pq` package downloads those zip files and writes their component
tables as Parquet files in a local `DATA_DIR` repository. It is based on the
workflow originally described in Ian Gow's note "Getting SEC EDGAR XBRL data,"
but packages the update logic and stores source-file metadata in each Parquet
file rather than in a separate sidecar table.

## Setup

SEC automated-access guidance asks users to provide a descriptive user agent
that includes contact information. `dera.pq` resolves this value from an
explicit argument, the `SEC_USER_AGENT` environment variable, or
`getOption("HTTPUserAgent")`.

```{r}
#| eval: false
library(dera.pq)

dera_set_user_agent("your_name@email_provider.com")
```

The package writes Parquet files under a local data repository identified by
`DATA_DIR`. You can set this explicitly:

```{r}
#| eval: false
Sys.setenv(DATA_DIR = "~/Dropbox/pq_data")
```

or let `dera_data_dir()` resolve or prompt for the directory:

```{r}
#| eval: false
dera_data_dir()
```

The resulting layout is:

```text
<DATA_DIR>/dera/
<DATA_DIR>/dera_notes/
```

## Listing available source files

The file-listing helpers scrape the SEC pages and issue `HEAD` requests for
each zip file so the returned table includes source `Last-Modified` metadata.
The `last_modified_utc` column is a canonical UTC representation used for
timezone-stable update checks.

```{r}
#| eval: false
available_dera()
available_dera_notes()
```

For code that parameterizes over the two data sets, use:

```{r}
#| eval: false
available_dera_files("dera")
available_dera_files("dera_notes")
```

## Updating all files

The high-level update functions compare SEC source metadata with metadata
embedded in the local Parquet files. A zip file is downloaded only when its SEC
`Last-Modified` timestamp is newer or when one or more expected Parquet files
are missing.

```{r}
#| eval: false
update_dera()
update_dera_notes()
```

This replaces the manual workflow of downloading a zip file, opening each
tab-delimited file, specifying column types, converting date fields, and writing
each table to Parquet. Those details are still important, but they now live in
the package table specifications and update functions rather than in an analysis
script.

`update_dera()` writes the standard Financial Statement Data Set tables:
`sub`, `tag`, `num`, and `pre`.

`update_dera_notes()` writes the Financial Statement and Notes tables:
`sub_notes`, `tag_notes`, `dim_notes`, `num_notes`, `txt_notes`, `ren_notes`,
`pre_notes`, and `cal_notes`.

## Updating one source file

The repository-level functions call lower-level functions that update a single
SEC zip file:

```{r}
#| eval: false
update_dera_file("2024q1.zip")
update_dera_notes_file("2024q1_notes.zip")
```

These are useful for testing a new period, repairing a missing period, or
building a small local sample.

## Embedded Parquet metadata

Each Parquet file written by `dera.pq` includes source metadata in the Parquet
schema metadata. The package stores the raw SEC header for auditability and a
parsed UTC timestamp for comparison.

```{r}
#| eval: false
path <- file.path(Sys.getenv("DATA_DIR"), "dera_notes", "sub_notes_2024q1.parquet")
dera_file_metadata(path)
```

Typical metadata fields include:

- `dera_dataset`
- `dera_source_file`
- `dera_source_url`
- `sec_last_modified`
- `last_modified`
- `last_modified_utc`

The UTC field avoids false update decisions when code is run in sessions with
different local time zones.

## Working with the Parquet files

Once created, the Parquet files can be read by Arrow, DuckDB, Polars, or other
tools. In R, Arrow can open all periods for a table with a wildcard-like file
listing:

```{r}
#| eval: false
library(arrow)
dplyr::open_dataset(
  list.files(
    file.path(Sys.getenv("DATA_DIR"), "dera_notes"),
    pattern = "^sub_notes_.*\\.parquet$",
    full.names = TRUE
  )
)
```

In Python Polars, the same repository can be scanned using `era_pl`:

```python
from era_pl import load_parquet

sub = load_parquet("sub_notes_*", "dera_notes")
```
